---
title: "Elexon data"
format: html
editor: source
---

```{r globalopt, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = FALSE, cache = F)
```

## packages

```{r include=FALSE}
require(tidyverse)
require(kableExtra)
require(parallel)
require(dplyr)
require(scales)
require(ggsci)

require(jsonlite)
require(httr)

require(data.table)
require(sf)
library(rnaturalearth)
library(rnaturalearthdata)

```

```{r}
# Set the theme for all plots
theme_set(theme_bw())
```

## Functions

```{r}
mc <- detectCores()-2
# list.files("~/OneDrive")
source("~/OneDrive/kablefunction.R")
```

## Data

```{r}
path <- "~/Documents/elexon"
elexon_path <- "/home/s2441782/Documents/elexon/"
```

### Aggregated Data

```{r}
# Actual Aggregated Generation Per Type (AGPT / B1620)

aggr.df <- read.csv(
  file.path(path,"aggregated_gen_type_20240718.csv")) %>% 
  mutate(StartTime = as.POSIXct(StartTime, format = "%Y-%m-%dT%H:%M:%OS"))

aggr.df %>% 
  # filter(grepl("Wind",PsrType, ignore.case = TRUE)  ) %>% 
  # filter(minute(StartTime)==0) %>%
  ggplot()+
  geom_line(aes(StartTime,Quantity,colour = PsrType, group = PsrType))+
  theme(legend.position = "bottom")+
  scale_x_datetime(date_labels = "%Y-%m-%d %H:%M", date_breaks = "1 day")

aggr.df %>% 
  filter(grepl("Wind|Solar",PsrType, ignore.case = TRUE)  ) %>%
  # filter(minute(StartTime)==0) %>%
  ggplot()+
  geom_line(aes(StartTime,Quantity,colour = PsrType, group = PsrType))+
  theme(legend.position = "bottom")+
  scale_x_datetime(date_labels = "%Y-%m-%d %H:%M", date_breaks = "1 day")+
  scale_color_lancet()
```

```{r}
aggr.df %>% 
  # filter(grepl("Wind",PsrType, ignore.case = TRUE)  ) %>% 
  select(StartTime,PsrType,Quantity)
```

```{r}
url <- 'https://data.elexon.co.uk/bmrs/api/v1/datasets/AGPT?publishDateTimeFrom=2024-07-18%2018%3A00&publishDateTimeTo=2024-07-21%2009%3A00&format=json' 
response <- GET(url, accept("text/plain"))
json_data <- content(response, "text", encoding = "UTF-8")

json_data %>% 
  fromJSON(flatten = T)
```

### forecast

```{r}
url <- "https://data.elexon.co.uk/bmrs/api/v1/datasets/DGWS/stream?publishDateTimeFrom=2024%2F07%2F18%2007%3A00&publishDateTimeTo=2024%2F07%2F21%2010%3A00"
response <- GET(url, accept("text/plain"))

json_data <- content(response, "text", encoding = "UTF-8")
forecast <- fromJSON(json_data) %>% 
  mutate(startTime = as.POSIXct(startTime, format = "%Y-%m-%dT%H:%M:%OS"))



```

```{r}
forecast %>% 
  ggplot()+
  geom_line(aes(startTime,quantity,colour = psrType, group = psrType))+
  theme(legend.position = "bottom")+
  scale_x_datetime(date_labels = "%Y-%m-%d %H:%M", date_breaks = "1 day")+
  scale_color_lancet()
```

### disaggregated data

#### list of BUs

```{r}
url <- "https://data.elexon.co.uk/bmrs/api/v1/reference/bmunits/all"
response <- GET(url, accept("text/plain"))
json_data <- content(response, "text", encoding = "UTF-8")
bmus <- fromJSON(json_data) 
```

```{r}
bmus %>% 
  filter(grepl("Hornsea",leadPartyName))
```

```{r}
# wind.bmus <- bmus %>% 
#   filter(grepl("WIND",fuelType), fpnFlag)
# wind.bmus
```

#### capacity

```{r}
url <- "https://data.elexon.co.uk/bmrs/api/v1/datasets/IGCPU?publishDateTimeFrom=2023-01-01T00%3A00%3A00Z&publishDateTimeTo=2024-01-02T00%3A00%3A00Z&format=json"

response <- GET(url, accept("text/plain"))
json_data <- content(response, "text", encoding = "UTF-8")
capacity <- fromJSON(json_data)$data
capacity %>% 
  filter(grepl("Wind|Solar",psrType, ignore.case = TRUE)  ) %>% 
  filter(!is.na(bmUnit)) %>% 
  arrange(desc(installedCapacity))
```
```{r}
capacity %>% 
  filter(grepl("Wind|Solar",psrType, ignore.case = TRUE)  ) %>%
  # group_by(psrType) %>% 
  summarise(n=n())
```

#### BMU subset

```{r}
url0 <- "https://data.elexon.co.uk/bmrs/api/v1/datasets/B1610/stream?from=2025-05-25T00%3A00Z&to=2025-06-04T00%3A00Z&bmUnit="
bmu.string <- "T_HOWBO-1&bmUnit=T_HOWBO-3&bmUnit=T_HOWBO-2&bmUnit=T_HOWAO-3&bmUnit=T_HOWAO-2&bmUnit=T_HOWAO-1"


url <- paste0(url0,bmu.string)

response <- GET(url, accept("text/plain"))
json_data <- content(response, "text", encoding = "UTF-8")
generation.bmu0 <- fromJSON(json_data)
```

```{r}
generation.bmu0 %>% 
  mutate(halfHourEndTime = as.POSIXct(halfHourEndTime, format = "%Y-%m-%dT%H:%M:%OS")) %>% 
  ggplot()+
  geom_line(aes(halfHourEndTime,quantity,colour = bmUnit, group = bmUnit))+
  theme(legend.position = "bottom")+
  scale_x_datetime(date_labels = "%Y-%m-%d %H:%M")+
  scale_color_lancet()+
  labs(y = "Generation (MW)", x = "time")
```

#### BMU full list

```{r}
url0 <- "https://data.elexon.co.uk/bmrs/api/v1/datasets/B1610/stream?from=2025-06-01T00%3A00Z&to=2025-06-04T00%3A00Z&bmUnit="

bmu.string <- capacity %>% 
  filter(grepl("Wind|Solar",psrType, ignore.case = TRUE)  ) %>% 
  filter(!is.na(bmUnit)) %>% 
  pull(bmUnit) %>% 
  paste(.,collapse = "&bmUnit=")


# bmu.string <- "T_HOWBO-1&bmUnit=T_HOWBO-3&bmUnit=T_HOWBO-2&bmUnit=T_HOWAO-3&bmUnit=T_HOWAO-2&bmUnit=T_HOWAO-1"


url <- paste0(url0,bmu.string)

response <- GET(url, accept("text/plain"))
json_data <- content(response, "text", encoding = "UTF-8")
generation.bmu <- fromJSON(json_data) %>% 
  mutate(halfHourEndTime = as.POSIXct(halfHourEndTime, format = "%Y-%m-%dT%H:%M:%OS")) 
```

```{r}
generation.bmu %>%
  pull(bmUnit) %>% 
  unique() %>% 
  length()
# view(generation.bmu)

generation.bmu %>% 
  group_by(halfHourEndTime,settlementPeriod) %>% 
  summarise(quantity = sum(quantity)) %>% 
  ggplot()+
  geom_line(aes(halfHourEndTime,quantity))+
  theme(legend.position = "bottom")+
  scale_x_datetime(date_labels = "%Y-%m-%d %H:%M", date_breaks = "1 day")+
  scale_color_lancet()
```

```{r}

aggr.df %>% 
  filter(grepl("Wind",PsrType, ignore.case = TRUE)  ) %>% 
  group_by(StartTime,SettlementPeriod) %>% 
  summarise(quantity = sum(Quantity)) %>% 
  mutate(dat.source="aggregate") %>% 
  bind_rows(
    generation.bmu %>% 
      group_by(halfHourEndTime,settlementPeriod) %>%
      summarise(quantity = sum(quantity)) %>% 
      mutate(StartTime=halfHourEndTime) %>% 
      mutate(dat.source="bmus") 
  ) %>% 
  ggplot()+
  geom_line(aes(StartTime,quantity,col=dat.source))+
  theme(legend.position = "bottom")+
  scale_x_datetime(date_labels = "%Y-%m-%d %H:%M", date_breaks = "1 day")+
  scale_color_lancet()
```

# updated code

### AGPT

```{r}
time_seq <- dates <- seq.Date(
  from = as.Date("2025-01-01"),
  to = as.Date("2025-05-26"),
  by = "4 days"
)

extraction <- lapply(
  seq_along(time_seq[-1]),
  
  
  \(i){
    
  t0 <- as.character(time_seq[i])
  t1 <- as.character(time_seq[i+1])
  dt0 <- paste(t0, "09:00")
  dt1 <- paste(t1, "09:00")
  dt0_enc <- URLencode(dt0, reserved = TRUE)
  dt1_enc <- URLencode(dt1, reserved = TRUE)
  url <- sprintf("https://data.elexon.co.uk/bmrs/api/v1/datasets/AGPT/stream?publishDateTimeFrom=%s&publishDateTimeTo=%s&format=json", dt0_enc, dt1_enc)
  
  response <- GET(url, accept("text/plain"))
  json_data <- content(response, "text", encoding = "UTF-8")
  # browser()
  agpt0 <- json_data %>% 
    fromJSON(flatten = T) %>% 
    mutate(across(matches("Time"), ~ as.POSIXct(., format = "%Y-%m-%dT%H:%M:%OS")))
  agpt0}
)



```

```{r}

agpt25 <- extraction %>% bind_rows()

# list.files(path)
agpt25 %>% 
  write.csv(
    gzfile(file.path(path,"aggregated_gen_type_2025.csv.gz"))
  )
```

####plots

```{r}
agpt25 %>% 
  filter(as.Date(settlementDate)>="2025-04-01") %>% 
  filter(grepl("Wind",psrType, ignore.case = TRUE)  ) %>%
  filter(minute(startTime)==0) %>%
  ggplot()+
  geom_line(aes(startTime,quantity,colour = psrType, group = psrType))+
  theme(legend.position = "bottom")+
  scale_x_datetime(date_labels = "%Y-%m-%d %H:%M")

agpt25 %>% 
  filter(as.Date(settlementDate)>="2025-05-15") %>% 
  filter(grepl("Wind|Solar",psrType, ignore.case = TRUE)  ) %>%
  # filter(minute(startTime)==0) %>%
  ggplot()+
  geom_line(aes(startTime,quantity,colour = psrType, group = psrType))+
  theme(legend.position = "bottom")+
  scale_x_datetime(date_labels = "%Y-%m-%d %H:%M")+
  scale_color_lancet()
```

### prediction

```{r}

time_seq <- dates <- seq.Date(
  from = as.Date("2025-01-01"),
  to = as.Date("2025-05-26"),
  by = "4 days"
)
base_url <- "https://data.elexon.co.uk/bmrs/api/v1/datasets/DGWS/stream?"

extraction <- lapply(
  seq_along(time_seq[-1]),
  
  
  \(i){
    
  t0 <- as.character(time_seq[i])
  t1 <- as.character(time_seq[i+1])
  dt0 <- paste(t0, "09:00")
  dt1 <- paste(t1, "09:00")
  dt0_enc <- URLencode(dt0, reserved = TRUE)
  dt1_enc <- URLencode(dt1, reserved = TRUE)
  url <- sprintf(
    "%spublishDateTimeFrom=%s&publishDateTimeTo=%s&format=json",
    base_url,dt0_enc, dt1_enc)
  
  response <- GET(url, accept("text/plain"))
  json_data <- content(response, "text", encoding = "UTF-8")
  # browser()
  df <- json_data %>% 
    fromJSON(flatten = T) %>% 
    mutate(across(matches("Time"), ~ as.POSIXct(., format = "%Y-%m-%dT%H:%M:%OS")))
  df}
)

dgws25 <- extraction %>% bind_rows()
dgws25 %>% 
  write.csv(
    gzfile(file.path(path,"dayahead_generation_windsolar_2025.csv.gz"))
  )
```

```{r}
dgws25 %>% 
  filter(processType %in% c("Day ahead")) %>% 
  filter(as.Date(settlementDate)>="2025-05-10") %>% 
  filter(minute(startTime)==0) %>%
  ggplot()+
  geom_line(aes(startTime,quantity,colour = psrType, group = psrType))+
  theme(legend.position = "bottom")+
  scale_x_datetime(date_labels = "%y-%m-%d %H:%M", breaks = "3 days")+
  scale_color_lancet()
```

### disaggregated

```{r}
url <- "https://data.elexon.co.uk/bmrs/api/v1/reference/bmunits/all"
response <- GET(url, accept("text/plain"))
json_data <- content(response, "text", encoding = "UTF-8")
bmus <- fromJSON(json_data) 
```

```{r}
bmus %>% 
  filter(grepl("Hornsea",leadPartyName))
```

```{r}
# wind.bmus <- bmus %>% 
#   filter(grepl("WIND",fuelType), fpnFlag) %>% 
#   filter(productionOrConsumptionFlag %in% c("P"))

wind.bmus <- bmus %>% 
  filter(
    grepl("WIND",fuelType) | 
      grepl("wind|offshore", leadPartyName, ignore.case = TRUE )|
      grepl("wind farm|windfarm", bmUnitName, ignore.case = TRUE )) %>% 
  filter(!is.na(elexonBmUnit))
wind.bmus
```

```{r}
wind.bmus %>% 
  write.csv(
    .,gzfile(file.path(path, "wind_bmu.csv.gz"))
  )
```

###capacity

```{r}



time_seq <- dates <- seq.Date(
  from = as.Date("2025-05-01"),
  to = as.Date("2025-05-26"),
  by = "4 days"
)

t0 <- "2023-01-01T"
t1 <- "2024-01-02T"
dt0 <- paste0(t0, "00:00")
dt1 <- paste0(t1, "00:00")
dt0_enc <- URLencode(dt0, reserved = TRUE)
dt1_enc <- URLencode(dt1, reserved = TRUE)
base_url <- "https://data.elexon.co.uk/bmrs/api/v1/datasets/IGCPU/stream?"
url <- sprintf(
  # "%spublishDateTimeFrom=2023-01-01T00%3A00%3A00Z&publishDateTimeTo=2024-01-02T00%3A00%3A00Z&format=json",
  "%spublishDateTimeFrom=%s&publishDateTimeTo=%s&format=json",
    base_url,dt0_enc, dt1_enc)

response <- GET(url, accept("text/plain"))
json_data <- content(response, "text", encoding = "UTF-8")
capacity <- fromJSON(json_data)
capacity %>% 
  filter(grepl("Wind|Solar",psrType, ignore.case = TRUE)  ) %>% 
  filter(!is.na(bmUnit)) %>% 
  arrange(desc(installedCapacity))

```
```{r}
capacity_latest <- capacity %>% 
  filter(!is.na(installedCapacity)) %>% 
  filter(installedCapacity>0) %>% 
  group_by(bmUnit, psrType) %>% 
  mutate(effectiveFrom = as.Date(effectiveFrom)) %>% 
  filter(effectiveFrom == max(effectiveFrom))
  
capacity_latest %>% 
  group_by(psrType) %>% 
  summarise(
    n = n(),installedCapacity = sum(installedCapacity, na.rm = T))
```

#### BMU full list

```{r}


t1 <- paste0("2025-06-01T","00:00Z") %>% URLencode(., reserved = TRUE)
t2 <- paste0("2025-06-04T","00:00Z") %>% URLencode(., reserved = TRUE)

base_url <- "https://data.elexon.co.uk/bmrs/api/v1/datasets/B1610/stream?"
url0 <- sprintf(
  "%sfrom=%s&to=%s&bmUnit=",
  base_url,t1,t2)

fragments <- 3
n_bmu <- nrow(wind.bmus)
parts_vec <- seq(1,n_bmu, length.out = fragments+1) %>% trunc()

# list strings in fragments
bmu_strings <- mapply(
  \(left,right){
    bmu.string <- wind.bmus %>%
      slice(left:right) %>%
      pull(elexonBmUnit) %>%
      paste(., collapse = "&bmUnit=")
    
    print(bmu.string)
  },
  left = parts_vec[1:fragments]+c(0,rep(1,fragments-1)),
  right = parts_vec[-1]
  )
# query data by fragment
bmu_df <- lapply(
  bmu_strings, 
  \(string){
    url <- paste0(url0,string)
    response <- GET(url, accept("text/plain"))
    json_data <- content(response, "text", encoding = "UTF-8")
    generation.bmu <- fromJSON(json_data) %>%
      mutate(halfHourEndTime = as.POSIXct(halfHourEndTime, format = "%Y-%m-%dT%H:%M:%OS")) 
    generation.bmu
  }) %>% bind_rows()
```

```{r}
bmu_df %>% #bind_rows() %>% 
  filter(grepl("HOWAO",bmUnit )) %>% 
  filter(as.Date(settlementDate)>="2025-06-01") %>% 
  filter(minute(halfHourEndTime)==0) %>%
  ggplot()+
  geom_line(aes(halfHourEndTime,quantity,colour = bmUnit))+
  theme(legend.position = "bottom")+
  scale_x_datetime(date_labels = "%y-%m-%d %H:%M")+
  scale_color_lancet()
```

```{r}

time_seq <- dates <- seq.Date(
  from = as.Date("2025-01-01"),
  to = as.Date("2025-06-01"),
  by = "4 days"
)
base_url <- "https://data.elexon.co.uk/bmrs/api/v1/datasets/B1610/stream?"

extraction <- lapply(
  seq_along(time_seq[-1]),
  
  \(i){
    
  t0 <- as.character(time_seq[i])
  t1 <- as.character(time_seq[i+1])
  dt0 <- paste(t0, "09:00")
  dt1 <- paste(t1, "09:00")
  dt0_enc <- URLencode(dt0, reserved = TRUE)
  dt1_enc <- URLencode(dt1, reserved = TRUE)
  url <- sprintf(
    "%sfrom=%s&to=%s&bmUnit=",
    base_url,dt0_enc, dt1_enc)
  # browser()
  fragments <- 3
  n_bmu <- nrow(wind.bmus)
  parts_vec <- seq(1,n_bmu, length.out = fragments+1) %>% trunc()
  
  # list strings in fragments
  bmu_strings <- mapply(
    \(left,right){
      bmu.string <- wind.bmus %>%
        slice(left:right) %>%
        pull(elexonBmUnit) %>%
        paste(., collapse = "&bmUnit=")
      
      # print(bmu.string)
    },
    left = parts_vec[1:fragments]+c(0,rep(1,fragments-1)),
    right = parts_vec[-1]
    )
  
  # query data by fragment
  bmu_df <- lapply(
    bmu_strings, 
    \(string){
      url <- paste0(url,string)
      response <- GET(url, accept("text/plain"))
      json_data <- content(response, "text", encoding = "UTF-8")
      generation.bmu <- fromJSON(json_data) %>%
        mutate(across(matches("Time"), ~ as.POSIXct(., format = "%Y-%m-%dT%H:%M:%OS")))
      generation.bmu
    }) %>% bind_rows()
  # browser()
  bmu_df
  }
)
bmu_25 <- extraction %>% bind_rows()

bmu_25 %>% 
  write.csv(
    gzfile(file.path(path,"wind_gen_bmu_2025.csv.gz"))
  )
```

```{r}
bmu_25 %>% 
  filter(grepl("HOWAO",bmUnit )) %>% 
  # filter(as.Date(settlementDate)>="2025-06-01") %>% 
  filter(minute(halfHourEndTime)==0) %>%
  ggplot()+
  geom_line(aes(halfHourEndTime,quantity,colour = bmUnit))+
  theme(legend.position = "bottom")+
  scale_x_datetime(date_labels = "%y-%m-%d %H:%M")+
  scale_color_lancet()
```

## Functions

### AGPT

```{r}
get_agpt <- function(
    year = 2024,
    path,
    file_name = sprintf("aggregated_gen_type_%d.csv.gz",year),
    end_time = NULL
){
  
  if (is.null(end_time)){
    time_seq <- dates <- seq.Date(
    from = as.Date(sprintf("%d-01-01",year)),
    to = as.Date(sprintf("%d-12-31",year)),
    by = "4 days"
    )
    time_seq <- c(time_seq, as.Date(sprintf("%d-01-01",year+1)))
  } else{
    time_seq <- dates <- seq.Date(
    from = as.Date(sprintf("%d-01-01",year)),
    to = as.Date(end_time),
    by = "4 days"
    )
    if(time_seq[length(time_seq)]<end_time) time_seq <- c(time_seq,end_time)
  }
  
  
  # browser()
  extraction <- lapply(
    seq_along(time_seq[-1]),
    
    
    \(i){
      
    t0 <- as.character(time_seq[i])
    t1 <- as.character(time_seq[i+1])
    dt0 <- paste(t0, "01:00")
    dt1 <- paste(t1, "01:00")
    dt0_enc <- URLencode(dt0, reserved = TRUE)
    dt1_enc <- URLencode(dt1, reserved = TRUE)
    
    base_url <- "https://data.elexon.co.uk/bmrs/api/v1/datasets/AGPT/stream?"
    
    url <- sprintf(
      "%spublishDateTimeFrom=%s&publishDateTimeTo=%s&format=json",
      base_url,dt0_enc, dt1_enc)
    
    response <- GET(url, accept("text/plain"))
    json_data <- content(response, "text", encoding = "UTF-8")
    # browser()
    agpt0 <- json_data %>% 
      fromJSON(flatten = T) %>% 
      mutate(across(matches("Time"), ~ as.POSIXct(., format = "%Y-%m-%dT%H:%M:%OS")))
    agpt0}
  )
  
  agptyear <- extraction %>% bind_rows()

  agptyear %>% 
    write.csv(
      gzfile(file.path(path,file_name))
    )
  
  invisible(agptyear)
}






agpt25 <- get_agpt(
  year = 2025,
  path = file.path(path,"data_by_year"),
  end_time = "2025-05-30")
```

```{r}
agpt24 <- get_agpt(
  year = 2024,
  path = file.path(path,"data_by_year"))
agpt23 <- get_agpt(
  year = 2023,
  path = file.path(path,"data_by_year"))

agpt22 <- get_agpt(
  year = 2022,
  path = file.path(path,"data_by_year"))

agpt21 <- get_agpt(
  year = 2021,
  path = file.path(path,"data_by_year"))

agpt20 <- get_agpt(
  year = 2020,
  path = file.path(path,"data_by_year"))

```

```{r}


```

### day ahead forecast

```{r}
time_seq <- dates <- seq.Date(
  from = as.Date("2025-01-01"),
  to = as.Date("2025-05-26"),
  by = "4 days"
)
base_url <- "https://data.elexon.co.uk/bmrs/api/v1/datasets/DGWS/stream?"

extraction <- lapply(
  seq_along(time_seq[-1]),
  
  
  \(i){
    
  t0 <- as.character(time_seq[i])
  t1 <- as.character(time_seq[i+1])
  dt0 <- paste(t0, "09:00")
  dt1 <- paste(t1, "09:00")
  dt0_enc <- URLencode(dt0, reserved = TRUE)
  dt1_enc <- URLencode(dt1, reserved = TRUE)
  url <- sprintf(
    "%spublishDateTimeFrom=%s&publishDateTimeTo=%s&format=json",
    base_url,dt0_enc, dt1_enc)
  
  response <- GET(url, accept("text/plain"))
  json_data <- content(response, "text", encoding = "UTF-8")
  # browser()
  df <- json_data %>% 
    fromJSON(flatten = T) %>% 
    mutate(across(matches("Time"), ~ as.POSIXct(., format = "%Y-%m-%dT%H:%M:%OS")))
  df}
)

dgws25 <- extraction %>% bind_rows()
dgws25 %>% 
  write.csv(
    gzfile(file.path(path,"dayahead_generation_windsolar_2025.csv.gz"))
  )

get_dwgs <- function(
    year = 2024,
    path,
    file_name = sprintf("dayahead_generation_windsolar_%d.csv.gz",year)
){
  
  time_seq <- dates <- seq.Date(
  from = as.Date(sprintf("%d-01-01",year)),
  to = as.Date(sprintf("%d-12-31",year)),
  by = "4 days"
  )
  
  # if(day(time_seq[length(time_seq)])!=31) 
    
  time_seq <- c(time_seq, as.Date(sprintf("%d-01-01",year+1)))
  
  # browser()
  extraction <- lapply(
    seq_along(time_seq[-1]),
    
    
    \(i){
      
    t0 <- as.character(time_seq[i])
    t1 <- as.character(time_seq[i+1])
    dt0 <- paste(t0, "01:00")
    dt1 <- paste(t1, "01:00")
    dt0_enc <- URLencode(dt0, reserved = TRUE)
    dt1_enc <- URLencode(dt1, reserved = TRUE)
    
    base_url <- "https://data.elexon.co.uk/bmrs/api/v1/datasets/DGWS/stream?"
    
    url <- sprintf(
      "%spublishDateTimeFrom=%s&publishDateTimeTo=%s&format=json",
      base_url,dt0_enc, dt1_enc)
    
    response <- GET(url, accept("text/plain"))
    json_data <- content(response, "text", encoding = "UTF-8")
    # browser()
    dwgs0 <- tryCatch(
      json_data %>% 
      fromJSON(flatten = T) #%>% 
      # mutate(across(matches("Time"), ~ as.POSIXct(., format = "%Y-%m-%dT%H:%M:%OS")))
      ,
      error = function(e) {
        warning(sprintf("Data extraction for day %s failed. Response: %s\n",t0,response$status_code))
        return(NULL)
        }
      )
    dwgs0}
  )
  
  dwgsyear <- extraction %>% bind_rows()

  dwgsyear %>% 
    write.csv(
      gzfile(file.path(path,file_name))
    )
  
  invisible(dwgsyear)
}

dwgs0 <- get_dwgs(
  year = 2022,
  path = file.path(path,"data_by_year"))
```

```{r}

lapply(
  2020:2021,
  \(y) get_dwgs(
         year = y,
         path = file.path(path,"data_by_year"))
       )

```

### BMU detail

```{r}
time_seq <- dates <- seq.Date(
  from = as.Date("2025-01-01"),
  to = as.Date("2025-06-01"),
  by = "4 days"
)
base_url <- "https://data.elexon.co.uk/bmrs/api/v1/datasets/B1610/stream?"

extraction <- lapply(
  seq_along(time_seq[-1]),
  
  \(i){
    
  t0 <- as.character(time_seq[i])
  t1 <- as.character(time_seq[i+1])
  dt0 <- paste(t0, "09:00")
  dt1 <- paste(t1, "09:00")
  dt0_enc <- URLencode(dt0, reserved = TRUE)
  dt1_enc <- URLencode(dt1, reserved = TRUE)
  url <- sprintf(
    "%sfrom=%s&to=%s&bmUnit=",
    base_url,dt0_enc, dt1_enc)
  # browser()
  fragments <- 3
  n_bmu <- nrow(wind.bmus)
  parts_vec <- seq(1,n_bmu, length.out = fragments+1) %>% trunc()
  
  # list strings in fragments
  bmu_strings <- mapply(
    \(left,right){
      bmu.string <- wind.bmus %>%
        slice(left:right) %>%
        pull(elexonBmUnit) %>%
        paste(., collapse = "&bmUnit=")
      
      # print(bmu.string)
    },
    left = parts_vec[1:fragments]+c(0,rep(1,fragments-1)),
    right = parts_vec[-1]
    )
  
  # query data by fragment
  bmu_df <- lapply(
    bmu_strings, 
    \(string){
      url <- paste0(url,string)
      response <- GET(url, accept("text/plain"))
      json_data <- content(response, "text", encoding = "UTF-8")
      generation.bmu <- fromJSON(json_data) %>%
        mutate(across(matches("Time"), ~ as.POSIXct(., format = "%Y-%m-%dT%H:%M:%OS")))
      generation.bmu
    }) %>% bind_rows()
  # browser()
  bmu_df
  }
)
bmu_25 <- extraction %>% bind_rows()

bmu_25 %>% 
  write.csv(
    gzfile(file.path(path,"wind_gen_bmu_2025.csv.gz"))
  )
```

```{r}
get_bmugen <- function(
    year = 2024,
    path,
    file_name = sprintf("wind_gen_bmu_%d.csv.gz",year),
    end_time = NULL
){
  
   if (is.null(end_time)){
    time_seq <- dates <- seq.Date(
    from = as.Date(sprintf("%d-01-01",year)),
    to = as.Date(sprintf("%d-12-31",year)),
    by = "4 days"
    )
    time_seq <- c(time_seq, as.Date(sprintf("%d-01-01",year+1)))
  } else{
    time_seq <- dates <- seq.Date(
    from = as.Date(sprintf("%d-01-01",year)),
    to = as.Date(end_time),
    by = "4 days"
    )
    if(time_seq[length(time_seq)]<end_time) time_seq <- c(time_seq,end_time)
  }
  
  # browser()
  extraction <- lapply(
    seq_along(time_seq[-1]),
    
    
    \(i){
      
    t0 <- as.character(time_seq[i])
    t1 <- as.character(time_seq[i+1])
    dt0 <- paste(t0, "01:00")
    dt1 <- paste(t1, "01:00")
    dt0_enc <- URLencode(dt0, reserved = TRUE)
    dt1_enc <- URLencode(dt1, reserved = TRUE)
    
    base_url <- "https://data.elexon.co.uk/bmrs/api/v1/datasets/B1610/stream?"
    
    url <- sprintf(
      "%sfrom=%s&to=%s&bmUnit=",
      base_url,dt0_enc, dt1_enc)
    
    fragments <- 3
    n_bmu <- nrow(wind.bmus)
    parts_vec <- seq(1,n_bmu, length.out = fragments+1) %>% trunc()
    
    # list strings in fragments
    bmu_strings <- mapply(
      \(left,right){
        bmu.string <- wind.bmus %>%
          slice(left:right) %>%
          pull(elexonBmUnit) %>%
          paste(., collapse = "&bmUnit=")
        
        # print(bmu.string)
      },
      left = parts_vec[1:fragments]+c(0,rep(1,fragments-1)),
      right = parts_vec[-1]
      )
    
    # query data by fragment
    bmu_df <- lapply(
      bmu_strings, 
      \(string){
        url <- paste0(url,string)
        response <- GET(url, accept("text/plain"))
        json_data <- content(response, "text", encoding = "UTF-8")
        generation.bmu <- tryCatch(
          fromJSON(json_data) %>%
          mutate(across(matches("Time"), ~ as.POSIXct(., format = "%Y-%m-%dT%H:%M:%OS"))),
        error = function(e) {
        warning(sprintf("Data extraction for day %s failed. Response: %s\n",t0,response$status_code))
        return(NULL)}
        )
      }) %>% bind_rows()
    
    bmu_df}
  )
  
  bmugenyear <- extraction %>% bind_rows()

  bmugenyear %>% 
    write.csv(
      gzfile(file.path(path,file_name))
    )
  
  invisible(bmugenyear)
}

bmugen0 <- get_bmugen(
  year = 2025,
  path = file.path(path,"data_by_year"),
  end_time = "2025-06-04 00:30:00 BST"
  )
```

```{r}
lapply(
  2020:2023,
  \(y) get_bmugen(
         year = y,
         path = file.path(path,"data_by_year"))
       )
```

###BOD

```{r}
get_bod <- function(
    year = 2024,
    path,
    file_name = sprintf("bid_offer_data_%d.csv.gz",year),
    end_time = NULL
){
  
   if (is.null(end_time)){
    time_seq <- dates <- seq.Date(
    from = as.Date(sprintf("%d-01-01",year)),
    to = as.Date(sprintf("%d-12-31",year)),
    by = "4 days"
    )
    time_seq <- c(time_seq, as.Date(sprintf("%d-01-01",year+1)))
  } else{
    time_seq <- dates <- seq.Date(
    from = as.Date(sprintf("%d-01-01",year)),
    to = as.Date(end_time),
    by = "4 days"
    )
    if(time_seq[length(time_seq)]<end_time) time_seq <- c(time_seq,end_time)
  }
  
  # browser()
  extraction <- lapply(
    seq_along(time_seq[-1]),
    
    
    \(i){
      
    t0 <- as.character(time_seq[i])
    t1 <- as.character(time_seq[i+1])
    dt0 <- paste(t0, "01:00")
    dt1 <- paste(t1, "01:00")
    dt0_enc <- URLencode(dt0, reserved = TRUE)
    dt1_enc <- URLencode(dt1, reserved = TRUE)
    
    base_url <- "https://data.elexon.co.uk/bmrs/api/v1/datasets/BOD/stream?"
    
    url <- sprintf(
      "%sfrom=%s&to=%s&bmUnit=",
      base_url,dt0_enc, dt1_enc)
    
    fragments <- 3
    n_bmu <- nrow(wind.bmus)
    parts_vec <- seq(1,n_bmu, length.out = fragments+1) %>% trunc()
    
    # list strings in fragments
    bmu_strings <- mapply(
      \(left,right){
        bmu.string <- wind.bmus %>%
          slice(left:right) %>%
          pull(elexonBmUnit) %>%
          paste(., collapse = "&bmUnit=")
        
        # print(bmu.string)
      },
      left = parts_vec[1:fragments]+c(0,rep(1,fragments-1)),
      right = parts_vec[-1]
      )
    
    # query data by fragment
    bmu_df <- lapply(
      bmu_strings, 
      \(string){
        url <- paste0(url,string)
        response <- GET(url, accept("text/plain"))
        json_data <- content(response, "text", encoding = "UTF-8")
        generation.bmu <- tryCatch(
          fromJSON(json_data) %>%
          mutate(across(matches("Time"), ~ as.POSIXct(., format = "%Y-%m-%dT%H:%M:%OS"))),
        error = function(e) {
        warning(sprintf("Data extraction for day %s failed. Response: %s\n",t0,response$status_code))
        return(NULL)}
        )
      }) %>% bind_rows()
    
    bmu_df}
  )
  
  bmugenyear <- extraction %>% bind_rows()

  bmugenyear %>% 
    write.csv(
      gzfile(file.path(path,file_name))
    )
  
  invisible(bmugenyear)
}
```

```{r}
bod_22 <- get_boa(
  year = 2022,
  path = file.path(path,"data_by_year")
)
```

```{r}
bod_22 %>% head()
bod_22 %>% 
  filter(grepl("HOWAO",bmUnit)) %>% 
  filter(settlementDate %in% "2022-09-26") %>% 
  view()
  
```

###BOA

```{r}

get_boa <- function(
    year = 2024,
    path,
    file_name = sprintf("bid_offer_accept_%d.csv.gz",year),
    end_time = NULL
){
  
   if (is.null(end_time)){
    time_seq <- dates <- seq.Date(
    from = as.Date(sprintf("%d-01-01",year)),
    to = as.Date(sprintf("%d-12-31",year)),
    by = "4 days"
    )
    time_seq <- c(time_seq, as.Date(sprintf("%d-01-01",year+1)))
  } else{
    time_seq <- dates <- seq.Date(
    from = as.Date(sprintf("%d-01-01",year)),
    to = as.Date(end_time),
    by = "4 days"
    )
    if(time_seq[length(time_seq)]<end_time) time_seq <- c(time_seq,end_time)
  }
  
  # browser()
  extraction <- lapply(
    seq_along(time_seq[-1]),
    
    
    \(i){
      
    t0 <- as.character(time_seq[i])
    t1 <- as.character(time_seq[i+1])
    dt0 <- paste(t0, "01:00")
    dt1 <- paste(t1, "01:00")
    dt0_enc <- URLencode(dt0, reserved = TRUE)
    dt1_enc <- URLencode(dt1, reserved = TRUE)
    
    base_url <- "https://data.elexon.co.uk/bmrs/api/v1/datasets/BOALF/stream?"
    
    url <- sprintf(
      "%sfrom=%s&to=%s&bmUnit=",
      base_url,dt0_enc, dt1_enc)
    
    fragments <- 3
    n_bmu <- nrow(wind.bmus)
    parts_vec <- seq(1,n_bmu, length.out = fragments+1) %>% trunc()
    
    # list strings in fragments
    bmu_strings <- mapply(
      \(left,right){
        bmu.string <- wind.bmus %>%
          slice(left:right) %>%
          pull(elexonBmUnit) %>%
          paste(., collapse = "&bmUnit=")
        
        # print(bmu.string)
      },
      left = parts_vec[1:fragments]+c(0,rep(1,fragments-1)),
      right = parts_vec[-1]
      )
    
    # query data by fragment
    bmu_df <- lapply(
      bmu_strings, 
      \(string){
        url <- paste0(url,string)
        response <- GET(url, accept("text/plain"))
        json_data <- content(response, "text", encoding = "UTF-8")
        generation.bmu <- tryCatch(
          fromJSON(json_data) %>%
          mutate(across(matches("Time"), ~ as.POSIXct(., format = "%Y-%m-%dT%H:%M:%OS"))),
        error = function(e) {
        warning(sprintf("Data extraction for day %s failed. Response: %s\n",t0,response$status_code))
        return(NULL)}
        )
      }) %>% bind_rows()
    
    bmu_df}
  )
  
  bmugenyear <- extraction %>% bind_rows()

  bmugenyear %>% 
    write.csv(
      gzfile(file.path(path,file_name))
    )
  
  invisible(bmugenyear)
}
```

```{r}
boa_22 <- get_boa(
  year = 2022,
  path = file.path(path,"data_by_year")
)
```

```{r}
# boa_22 <- fread(file.path(path,"data_by_year",))
boa_22 %>% 
  filter(grepl("HOWAO-1",bmUnit)) %>% 
  filter(settlementDate %in% "2022-09-26") %>% 
  arrange(settlementPeriodFrom, settlementPeriodTo) %>% 
  # group_by(settlementPeriodFrom,settlementPeriodTo) %>%
  # summarise()
  view()


```

```{r}
boa_22 %>% 
  filter(grepl("HOWAO",bmUnit)) %>% 
  filter(settlementDate %in% "2022-09-26") %>% 
  # filter(timeFrom > "2022-09-26 03:00", timeTo > "2022-09-26 06:00") %>% 
  arrange(settlementPeriodFrom, settlementPeriodTo) %>% 
  view()
  select(settlementPeriodFrom:acceptanceTime)
```

```{r}
boa_22 %>% 
  filter(grepl("HOWAO-1",bmUnit)) %>% 
  filter(settlementDate %in% "2022-09-26") %>% 
  arrange(settlementPeriodFrom, settlementPeriodTo) %>% 
  select(settlementPeriodFrom:acceptanceTime)
```

### PN
```{r}
get_pn <- function(
    year = 2024,
    path,
    file_name = sprintf("phys_notif_%d.csv.gz",year),
    end_time = NULL
){
  
   if (is.null(end_time)){
    time_seq <- dates <- seq.Date(
    from = as.Date(sprintf("%d-01-01",year)),
    to = as.Date(sprintf("%d-12-31",year)),
    by = "4 days"
    )
    time_seq <- c(time_seq, as.Date(sprintf("%d-01-01",year+1)))
  } else{
    time_seq <- dates <- seq.Date(
    from = as.Date(sprintf("%d-01-01",year)),
    to = as.Date(end_time),
    by = "4 days"
    )
    if(time_seq[length(time_seq)]<end_time) time_seq <- c(time_seq,end_time)
  }
  
  # browser()
  extraction <- lapply(
    seq_along(time_seq[-1]),
    
    
    \(i){
      
    t0 <- as.character(time_seq[i])
    t1 <- as.character(time_seq[i+1])
    dt0 <- paste(t0, "01:00")
    dt1 <- paste(t1, "01:00")
    dt0_enc <- URLencode(dt0, reserved = TRUE)
    dt1_enc <- URLencode(dt1, reserved = TRUE)
    
    base_url <- "https://data.elexon.co.uk/bmrs/api/v1/datasets/PN/stream?"
    
    url <- sprintf(
      "%sfrom=%s&to=%s&bmUnit=",
      base_url,dt0_enc, dt1_enc)
    
    fragments <- 3
    n_bmu <- nrow(wind.bmus)
    parts_vec <- seq(1,n_bmu, length.out = fragments+1) %>% trunc()
    
    # list strings in fragments
    bmu_strings <- mapply(
      \(left,right){
        bmu.string <- wind.bmus %>%
          slice(left:right) %>%
          pull(elexonBmUnit) %>%
          paste(., collapse = "&bmUnit=")
        
        # print(bmu.string)
      },
      left = parts_vec[1:fragments]+c(0,rep(1,fragments-1)),
      right = parts_vec[-1]
      )
    
    # query data by fragment
    bmu_df <- lapply(
      bmu_strings, 
      \(string){
        url <- paste0(url,string)
        response <- GET(url, accept("text/plain"))
        json_data <- content(response, "text", encoding = "UTF-8")
        generation.bmu <- tryCatch(
          fromJSON(json_data) %>%
          mutate(across(matches("Time"), ~ as.POSIXct(., format = "%Y-%m-%dT%H:%M:%OS"))),
        error = function(e) {
        warning(sprintf("Data extraction for day %s failed. Response: %s\n",t0,response$status_code))
        return(NULL)}
        )
      }) %>% bind_rows()
    
    bmu_df}
  )
  
  bmugenyear <- extraction %>% bind_rows()

  bmugenyear %>% 
    write.csv(
      gzfile(file.path(path,file_name))
    )
  
  invisible(bmugenyear)
}
```

```{r}
pn_24 <- get_pn(
  year = 2024,
  path = file.path(path,"data_by_year")
)
```

```{r}
# pn_24 <- boa_24
pn_24 %>% 
  filter(settlementDate %in% c("2024-12-18")) %>% 
  group_by(timeTo) %>% 
  summarise(
    levelTo = sum(levelTo),
    levelFrom = sum(levelFrom)) %>% 
  view()
```

### QPN

## Figures

### total vs day-ahead

```{r}
# require(data.table)
# read 2025 agpt
elexon_path <- "/home/s2441782/Documents/elexon/"
# list.files(file.path(elexon_path,"data_by_year"))
agpt <- fread(
  file.path(
    elexon_path,
    "data_by_year",
    "aggregated_gen_type_2025.csv.gz"))

# read 2025 forecast
dgws <- fread(
  file.path(
    elexon_path,
    "data_by_year",
    "dayahead_generation_windsolar_2025.csv.gz"))

```

```{r}
dgws
```

```{r}
agpt$psrType %>% unique() %>% paste(., collapse = ", ")
```

```{r}
agpt %>% 
  # filter(as.Date(settlementDate)>="2025-04-01") %>% 
  filter(grepl("Wind",psrType, ignore.case = TRUE)  ) %>%
  filter(minute(startTime)==0) %>%
  ggplot()+
  geom_line(aes(startTime,quantity,colour = psrType, group = psrType))+
  theme(legend.position = "bottom")+
  scale_x_datetime(date_labels = "%Y-%m-%d %H:%M")+
  scale_y_continuous(limits = c(0,12500))+
  scale_color_aaas()


agpt %>% 
  group_by(settlementDate, psrType) %>% 
  summarise(quantity = mean(quantity), .groups = "drop") %>% 
  filter(grepl("Wind",psrType, ignore.case = TRUE)  ) %>%
  ggplot()+
  geom_line(aes(settlementDate,quantity,colour = psrType, group = psrType))+
  theme(legend.position = "bottom")+
  scale_x_date()+
  scale_y_continuous(limits = c(0,12500))+
  scale_color_aaas()
  # scale_x_datetime(date_labels = "%Y-%m-%d %H:%M")
```

```{r}
dwgs %>% 
  filter(processType %in% c("Day ahead")) %>% 
  # filter(as.Date(settlementDate)>="2025-04-01") %>% 
  filter(grepl("Wind",psrType, ignore.case = TRUE)  ) %>%
  filter(minute(startTime)==0) %>%
  ggplot()+
  geom_line(aes(startTime,quantity,colour = psrType, group = psrType))+
  theme(legend.position = "bottom")+
  scale_x_datetime(date_labels = "%Y-%m-%d %H:%M")+
  scale_y_continuous(limits = c(0,12500))+
  scale_color_aaas()

dwgs %>%
  filter(processType %in% c("Day ahead")) %>% 
  group_by(settlementDate, psrType) %>% 
  summarise(quantity = mean(quantity,na.rm = T), .groups = "drop") %>% 
  filter(grepl("Wind",psrType, ignore.case = TRUE)  ) %>%
  ggplot()+
  geom_line(aes(settlementDate,quantity,colour = psrType, group = psrType))+
  theme(legend.position = "bottom")+
  scale_x_date()+
  scale_y_continuous(limits = c(0,12500))+
  scale_color_aaas()
```

### aggregation vs totals

```{r, fig.height=7, fig.width=6}

dwgs %>%
  filter(processType %in% c("Day ahead")) %>% 
  group_by(settlementDate, psrType) %>% 
  summarise(quantity = mean(quantity,na.rm = T), .groups = "drop") %>% 
  mutate(series = "forecast") %>% 
  bind_rows(
    agpt %>% 
      group_by(settlementDate, psrType) %>% 
      summarise(quantity = mean(quantity), .groups = "drop") %>% 
      filter(grepl("Wind|Solar",psrType, ignore.case = TRUE)  ) %>%
      mutate(series = "observed")
  ) %>% 
  mutate(
    psrType = factor(
      psrType, 
      levels = c("Wind Offshore","Wind Onshore","Solar")),
    series = factor(series, levels = c("observed","forecast"))
    ) %>% 
  ggplot()+
  geom_line(aes(settlementDate,quantity,colour = series))+
  facet_wrap(~psrType,nrow = 3)+
  theme(legend.position = "bottom")+
  scale_x_date()+
  scale_y_continuous(limits = c(0,12500))+
  scale_color_manual(values = c(pal_aaas()(2)[2],pal_aaas()(2)[1]))+
  labs(y="generation (MW)", x = "Date", col = "")

```

### BMU ids

```{r}
# list.files(path)
id_catalog <- readxl::read_xlsx(
  path = file.path(path,"osuked","id_dictionary.xlsx")
) %>% 
  drop_na() %>% 
   rename(
    rep_id = `REPD ID (New)`,
    bmu_id = `Settlement BMU ID`,
    dict_id = `Dictionary ID`,
    name = `Common Name`
    )

repdb <- fread(
  file.path(path,"osuked","renewable-energy-planning-database-q2-june-2021.csv")
)

repdb %>% 
  filter(grepl("Wind|Solar",`Technology Type`)) %>% 
  filter(`Development Status (short)` %in% c("Operational")) %>% 
  group_by(`Technology Type`) %>% 
  summarise(
    n = n(),
    capacity_available = sum(!is.na(`CfD Capacity (MW)`)),
    location_available = sum(!is.na(`X-coordinate`))) %>% 
  janitor::adorn_totals() %>% 
  kable() %>% 
  kable_styling(full_width = F)
```

### map

#### Renewable energy planning database (UK gov)

```{r fig.width=5, fig.height=7}


uk_ie <- ne_countries(scale = "medium", returnclass = "sf") %>%
  filter(admin %in% c("United Kingdom", "Ireland")) %>%
  st_transform(crs = 4326)


repdb %>% 
  rename(
    tech_typ = `Technology Type`,
    devel_stat = `Development Status (short)`,
    x_coord = `X-coordinate`,
    y_coord = `Y-coordinate`) %>% 
  filter(grepl("Wind|Solar",tech_typ)) %>% 
  filter(devel_stat %in% c("Operational")) %>% 
  filter(!is.na(x_coord)) %>% 
  st_as_sf(coords = c("x_coord","y_coord"), crs = 27700) %>% 
   st_transform(4326) %>% 
  ggplot()+
  geom_sf(data = uk_ie, fill = "gray95", color = "gray70") +  # map background
  geom_sf(aes(geometry = geometry, col = tech_typ),size = 1)+
  coord_sf(datum = 4326, crs = 4326)+
  scale_color_manual(values = pal_aaas()(3))+
  labs(col = "")+ theme(legend.position = "bottom")
```

#### map for BMUs

```{r}
# list.files(file.path(path,"data_by_year"))
# read bmu detal
bmu_df <- fread(
  file.path(path,"data_by_year","wind_gen_bmu_2025.csv.gz")
)
bmu_df$bmUnit %>% unique()
```

#### finding elexon BMUs
too restrictive
```{r fig.width=5, fig.height=7}
repd_subset <- repdb %>% 
  rename(
    ref_id = `Ref ID`,
    tech_typ = `Technology Type`,
    devel_stat = `Development Status (short)`,
    x_coord = `X-coordinate`,
    y_coord = `Y-coordinate`) %>% 
  filter(grepl("Wind|Solar",tech_typ)) %>% 
  filter(devel_stat %in% c("Operational")) 

repd_wind <- repdb %>% 
  rename(
    ref_id = `Ref ID`,
    tech_typ = `Technology Type`,
    devel_stat = `Development Status (short)`,
    x_coord = `X-coordinate`,
    y_coord = `Y-coordinate`) %>% 
  filter(grepl("Wind",tech_typ))# %>% 
  # filter(devel_stat %in% c("Operational")) 

# repd_subset$ref_id %>%uniqu %>%  length()

bmu_in_cat <- id_catalog %>% 
  filter(rep_id != "-") %>% 
  pull(bmu_id) %>% 
  str_split(., ",\\s*") %>% 
  unlist()

repdid_in_cat <- id_catalog %>% 
  filter(rep_id != "-") %>% 
  pull(rep_id) %>% 
  str_split(., ",\\s*") %>% 
  unlist()

repd_subset %>% 
  filter(ref_id %in% repdid_in_cat) %>%
  st_as_sf(coords = c("x_coord","y_coord"), crs = 27700) %>% 
   st_transform(4326) %>% 
  ggplot()+
  geom_sf(data = uk_ie, fill = "gray95", color = "gray70") +  # map background
  geom_sf(aes(geometry = geometry, col = tech_typ),size = 1)+
  coord_sf(datum = 4326, crs = 4326)+
  # scale_color_aaas()+
  scale_color_manual(values = pal_aaas()(3)[c(2,3)])+
  labs(col = "")+ theme(legend.position = "bottom")
```

```{r}
list.files(file.path(path))
loc_alt <- read.csv(file.path(path,"plant-locations.csv"))
id_catalog %>% 
  separate_rows(bmu_id, sep = ",\\s*")

loc_alt %>% 
  filter(dictionary_id %in% id_catalog$dict_id[])
  

elexon_bmus <- wind.bmus$elexonBmUnit %>% unique()

# sum(elexon_bmus %in% bmu_in_cat )
```

#### extra elexon bmus

```{r fig.height=7, fig.width=5}
# loc_alt %>% 
#   filter()


bmu_to_dict <- id_catalog %>% 
  separate_rows(bmu_id, sep = ",\\s*") %>%
  select(dict_id, name, bmu_id, rep_id)

bmu_25 <- bmu_df %>% 
  group_by(bmUnit) %>% 
  summarise(quantity = mean(quantity, na.rm = TRUE)) %>% 
  left_join(
    bmu_to_dict,
    by = c("bmUnit"= "bmu_id")
  ) %>% 
  left_join(
    loc_alt,
    by = c("dict_id"= "dictionary_id")
  ) %>% 
  filter(!is.na(longitude)) %>%
  st_as_sf(coords = c("longitude","latitude"), crs = 4326)

bmu_25 %>% 
  ggplot()+
  geom_sf(data = uk_ie, fill = "gray95", color = "gray70") +  # map background
  # geom_sf(aes(geometry = geometry, size = quantity), col = "darkgreen")+
  geom_sf(aes(geometry = geometry), col =  pal_aaas()(3)[c(3)])+
  labs(col ="Mean day generation MW")+
  theme(legend.position = "bottom")

```

####completing dictionary

```{r}
library(stringr)
clean_names <- function(name) {
  name %>%
    iconv(from = "", to = "UTF-8", sub = "") %>%     # Fix encoding
    tolower() %>%                                    # Lowercase
    str_replace_all("[^a-z0-9 ]", " ") %>%           # Remove punctuation
    str_replace_all("\\s+", " ") %>%                 # Remove extra spaces
    str_trim()
}
repdb <- readxl::read_xlsx(
  file.path(path,"renew_energ_plan_db","repd-q1-apr-2025.xlsx"),
  sheet = 2
) %>% 
  rename(
    ref_id = `Ref ID`,
    tech_typ = `Technology Type`,
    devel_stat = `Development Status (short)`,
    x_coord = `X-coordinate`,
    y_coord = `Y-coordinate`) %>% 
  mutate(clean_name = ifelse(is.na(clean_name),clean_names(`Site Name`),clean_name)) %>% 
  filter(!is.na(clean_name))


# info through robin hawkes website

bmu_catalog <- wind.bmus %>% 
  left_join(
      bmu_to_dict,
      by = c("elexonBmUnit"= "bmu_id")
    ) %>% 
  left_join(
    loc_alt,
    by = c("dict_id"= "dictionary_id")
  ) %>% 
  mutate(
    clean_name = clean_names(leadPartyName),
    clean_name = case_when(
      clean_name %in% c(
        "edf energy customers limited",
        "edf energy customers limited",
        "eneco energy trade",
        "erova energy limited",
        "smartestenergy limited",
        "statkraft markets gmbh",
        "npower commercial gas limited",
        "vattenfall energy trading") ~ clean_names(bmUnitName),
      
      TRUE ~ clean_name
    ))
# repdb through id_catalog


bmu_catalog2 <- bmu_catalog %>% 
  left_join(
    repdb %>% 
      select(clean_name,`Site Name`,ref_id, tech_typ,
             `Installed Capacity (MWelec)`,`Turbine Capacity`,`No. of Turbines`,
             devel_stat,Region,x_coord,y_coord,),
    by = c("clean_name")
  )


missing_latlon <- bmu_catalog2 %>%
  filter(is.na(latitude) & !is.na(x_coord) & !is.na(y_coord)) %>%
  st_as_sf(coords = c("x_coord", "y_coord"), crs = 27700) %>%
  st_transform(4326)

missing_latlon_coords <- missing_latlon %>%
  mutate(
    longitude = st_coordinates(.)[,1],
    latitude = st_coordinates(.)[,2]
  ) %>%
  st_drop_geometry()


bmu_catalog2 <- bmu_catalog2 %>% 
  filter(!is.na(latitude)) %>% 
  bind_rows(
    missing_latlon_coords
  )
```
```{r}
bmu_25 <- bmu_df %>% 
  group_by(bmUnit) %>% 
  summarise(quantity = mean(quantity, na.rm = TRUE)) %>% 
  left_join(
    bmu_catalog2,
    by = c("bmUnit" = "elexonBmUnit")
  ) 


bmu_25_sf <- bmu_25 %>% 
  filter(!is.na(longitude)) %>% 
   st_as_sf(coords = c("longitude","latitude"), crs = 4326) 

mpow25 <- bmu_25 %>% 
  summarise(quantity = sum(quantity)) %>% pull(quantity)

mpow25_loc <- bmu_25_sf %>% 
  summarise(quantity = sum(quantity)) %>% pull(quantity)

mpow25_loc/mpow25

p <- bmu_25_sf %>%  
  ggplot()+
  geom_sf(data = uk_ie, fill = "gray95", color = "gray70") +  # map background
  geom_sf(aes(geometry = geometry, col = quantity, text = paste("BMU Name:", bmUnitName)), size = 1)+
  coord_sf(datum = 4326, crs = 4326)+
  # scale_color_aaas()+
  # scale_color_manual(values = pal_aaas()(3)[c(2,3)])+
  labs(col = "mean generation MW", title = "Wind BM Units 2025")+ theme(legend.position = "bottom")+
  scale_color_viridis_c()
p
# plotly::ggplotly(
#   p
# )

```

### BMU data check

#### updated coverage with MWh conversion

##### 2025

```{r}
bmu_df <- fread(
  file.path(path,"data_by_year","wind_gen_bmu_2025.csv.gz")
)

elex_cov <- bmu_df %>% 
  # aggregate half hour generation
  group_by(bmUnit) %>% arrange(halfHourEndTime) %>% 
  mutate(quantity = quantity + lag(quantity)) %>% 
  filter(quantity>0) %>% 
  filter(minute(halfHourEndTime) == 0) %>% 
  group_by(
    halfHourEndTime
  ) %>% summarise(bmu_total = sum(quantity, na.rm = T)) %>% 
  left_join(
    agpt %>% 
      filter(grepl("Wind",psrType, ignore.case = TRUE)  ) %>%
      filter(minute(startTime)==0) %>% 
      group_by(startTime) %>% summarise(total = sum(quantity, na.rm = T)),
    by = c("halfHourEndTime" = "startTime")
  ) 

elex_cov %>% 
  mutate(date = date(halfHourEndTime)) %>% 
  group_by(date) %>% 
  summarise(across(matches("total"),~mean(., na.rm = TRUE))) %>%
  pivot_longer(cols = matches("total")) %>% 
  ggplot()+
  geom_line(aes(date,value,colour = name))+
  theme(legend.position = "bottom")+
  scale_x_date()+
  scale_y_continuous(limits = c(0,20000))+
  scale_color_manual(values = c(pal_aaas()(2)[2],pal_aaas()(2)[1]), labels = c("BMU","AGPT"))+
  labs(y="generation (MW)", x = "Date", col = "")
```

```{r}
bmu_df %>% 
  pull(bmUnit) %>% unique() %>% length()
elex_cov %>% 
  mutate(date = date(halfHourEndTime)) %>% 
  group_by(date) %>% 
  summarise(across(matches("total"),~mean(., na.rm = TRUE))) %>% 
  mutate(perc = bmu_total/total*100) %>% 
  summarise(coverage = mean(perc, na.rm =T))
```

##### 2024

```{r}
bmu_df <- fread(
  file.path(path,"data_by_year","wind_gen_bmu_2024.csv.gz")
)

agpt <- fread(
  file.path(
    elexon_path,
    "data_by_year",
    "aggregated_gen_type_2024.csv.gz"))

elex_cov <- bmu_df %>% 
  # aggregate half hour generation
  group_by(bmUnit) %>% arrange(halfHourEndTime) %>% 
  mutate(quantity = quantity + lag(quantity)) %>% 
  filter(quantity>0) %>% 
  filter(minute(halfHourEndTime) == 0) %>% 
  group_by(
    halfHourEndTime
  ) %>% summarise(bmu_total = sum(quantity, na.rm = T)) %>% 
  left_join(
    agpt %>% 
      filter(grepl("Wind",psrType, ignore.case = TRUE)  ) %>%
      filter(minute(startTime)==0) %>% 
      group_by(startTime) %>% summarise(total = sum(quantity, na.rm = T)),
    by = c("halfHourEndTime" = "startTime")
  ) 

elex_cov %>% 
  mutate(date = date(halfHourEndTime)) %>% 
  group_by(date) %>% 
  summarise(across(matches("total"),~mean(., na.rm = TRUE))) %>%
  pivot_longer(cols = matches("total")) %>% 
  ggplot()+
  geom_line(aes(date,value,colour = name))+
  theme(legend.position = "bottom")+
  scale_x_date()+
  scale_y_continuous(limits = c(0,20000))+
  scale_color_manual(values = c(pal_aaas()(2)[2],pal_aaas()(2)[1]))+
  labs(y="generation (MW)", x = "Date", col = "")
```

```{r}
bmu_df %>% 
  pull(bmUnit) %>% unique() %>% length()
elex_cov %>% 
  mutate(date = date(halfHourEndTime)) %>% 
  group_by(date) %>% 
  summarise(across(matches("total"),~mean(., na.rm = TRUE))) %>% 
  mutate(perc = bmu_total/total*100) %>% 
  summarise(coverage = mean(perc, na.rm =T))
```

```{r}
# agpt 2024
agpt24_wind <- agpt %>% 
  filter(grepl("Wind",psrType)) %>%
  group_by(startTime) %>% filter(publishTime == max(publishTime)) %>% ungroup() %>%
  group_by(startTime) %>%
  summarise(quantity = sum(quantity)) %>% 
  arrange(startTime) 

agpt24_wind %>% 
  ggplot()+
  geom_line(aes(startTime,quantity))


# agpt 2024
agpt24_wind2 <- agpt %>% 
  filter(grepl("Wind",psrType)) %>%
  group_by(startTime) %>% filter(publishTime == max(publishTime)) %>% ungroup() %>%
  # group_by(startTime) %>%
  # summarise(quantity = sum(quantity)) %>% 
  arrange(startTime) 

agpt24_wind2 %>% 
  ggplot()+
  geom_line(aes(startTime,quantity, col = psrType))

# agpt 2024
agpt %>% 
  filter(!grepl("Wind",psrType)) %>%
  group_by(startTime) %>% filter(publishTime == max(publishTime)) %>% ungroup() %>%
  arrange(startTime) %>% 
  group_by(psrType,settlementDate) %>% summarise(quantity = mean(quantity, na.rm = TRUE)) %>% 
  ggplot()+
  geom_line(aes(settlementDate,quantity, col = psrType))

```

######vs Day ahead

```{r}

# read 2025 forecast
dwgs <- fread(
  file.path(
    elexon_path,
    "data_by_year",
    "dayahead_generation_windsolar_2024.csv.gz"))


dwgs %>% 
  filter(processType %in% c("Day ahead")) %>% 
  group_by(startTime) %>% filter(publishTime == max(publishTime)) %>% ungroup() %>%
  filter(grepl("Wind",psrType, ignore.case = TRUE)  ) %>%
  filter(minute(startTime)==0) %>%
  ggplot()+
  geom_line(aes(startTime,quantity,colour = psrType, group = psrType))+
  theme(legend.position = "bottom")+
  scale_x_datetime(date_labels = "%Y-%m-%d %H:%M")+
  scale_y_continuous(limits = c(0,12500))+
  scale_color_aaas()

dwgs %>%
  filter(processType %in% c("Day ahead")) %>% 
  group_by(startTime) %>% filter(publishTime == max(publishTime)) %>% ungroup() %>%
  group_by(settlementDate, psrType) %>% 
  summarise(quantity = mean(quantity,na.rm = T), .groups = "drop") %>% 
  filter(grepl("Wind",psrType, ignore.case = TRUE)  ) %>%
  ggplot()+
  geom_line(aes(settlementDate,quantity,colour = psrType, group = psrType))+
  theme(legend.position = "bottom")+
  scale_x_date()+
  scale_y_continuous(limits = c(0,12500))+
  scale_color_aaas()

```

```{r}

dwgs %>%
  filter(processType %in% c("Day ahead")) %>% 
  group_by(settlementDate, psrType) %>% 
  summarise(quantity = mean(quantity,na.rm = T), .groups = "drop") %>% 
  mutate(series = "forecast") %>% 
  bind_rows(
    agpt %>% 
      group_by(settlementDate, psrType) %>% 
      summarise(quantity = mean(quantity), .groups = "drop") %>% 
      filter(grepl("Wind|Solar",psrType, ignore.case = TRUE)  ) %>%
      mutate(series = "observed")
  ) %>% 
  mutate(
    psrType = factor(
      psrType, 
      levels = c("Wind Offshore","Wind Onshore","Solar")),
    series = factor(series, levels = c("observed","forecast"))
    ) %>% 
  ggplot()+
  geom_line(aes(settlementDate,quantity,colour = series))+
  facet_wrap(~psrType,nrow = 3)+
  theme(legend.position = "bottom")+
  scale_x_date()+
  scale_y_continuous(limits = c(0,12500))+
  scale_color_manual(values = c(pal_aaas()(2)[2],pal_aaas()(2)[1]))+
  labs(y="generation (MW)", x = "Date", col = "")
```

##### 2023

```{r}
bmu_df <- fread(
  file.path(path,"data_by_year","wind_gen_bmu_2023.csv.gz")
)

agpt <- fread(
  file.path(
    elexon_path,
    "data_by_year",
    "aggregated_gen_type_2023.csv.gz"))

elex_cov <- bmu_df %>% 
  # aggregate half hour generation
  group_by(bmUnit) %>% arrange(halfHourEndTime) %>% 
  mutate(quantity = quantity + lag(quantity)) %>% 
  filter(quantity>0) %>% 
  filter(minute(halfHourEndTime) == 0) %>% 
  group_by(
    halfHourEndTime
  ) %>% summarise(bmu_total = sum(quantity, na.rm = T)) %>% 
  left_join(
    agpt %>% 
      filter(grepl("Wind",psrType, ignore.case = TRUE)  ) %>%
      filter(minute(startTime)==0) %>% 
      group_by(startTime) %>% summarise(total = sum(quantity, na.rm = T)),
    by = c("halfHourEndTime" = "startTime")
  ) 

elex_cov %>% 
  mutate(date = date(halfHourEndTime)) %>% 
  group_by(date) %>% 
  summarise(across(matches("total"),~mean(., na.rm = TRUE))) %>%
  pivot_longer(cols = matches("total")) %>% 
  ggplot()+
  geom_line(aes(date,value,colour = name))+
  theme(legend.position = "bottom")+
  scale_x_date()+
  scale_y_continuous(limits = c(0,20000))+
  scale_color_manual(values = c(pal_aaas()(2)[2],pal_aaas()(2)[1]))+
  labs(y="generation (MW)", x = "Date", col = "")
```

```{r}
bmu_df %>% 
  pull(bmUnit) %>% unique() %>% length()
elex_cov %>% 
  mutate(date = date(halfHourEndTime)) %>% 
  group_by(date) %>% 
  summarise(across(matches("total"),~mean(., na.rm = TRUE))) %>% 
  mutate(perc = bmu_total/total*100) %>% 
  summarise(coverage = mean(perc, na.rm =T))
```

# BMU CHECK

### hornsea 2022

BMU data from B1610 is MWh, multiply by 2 or add two half hour periods to get hourly generation. its missing bid offer acceptances.

```{r}
bmu_22 <-  fread(
  file.path(path,"data_by_year","wind_gen_bmu_2022.csv.gz")
)

bmu_22 %>% 
  filter(grepl("HOWAO", bmUnit)) %>% 
  filter(settlementDate %in% c(as.Date("2022-09-26"))) %>% 
  group_by(halfHourEndTime) %>% summarise(quantity = sum(quantity), .groups = "drop") %>% 
  arrange(halfHourEndTime) %>% 
  mutate(
    q2 = quantity*2,
    qh = quantity +lag(quantity)) %>% 
  view()
```

```{r}
# date range
bmu_df$settlementDate %>% range()


repd_subset


```

```{r}
bmu_t0 <- bmu_df %>% 
  select(-c(V1:psrType)) %>% 
  filter(
    settlementDate %in% c("2025-05-29")
  ) %>% 
  filter(
    halfHourEndTime > as.POSIXct("2025-05-29 05:00:00", tz="UTC"),
    halfHourEndTime <= as.POSIXct("2025-05-29 10:00:00", tz="UTC"),
    ) %>% 
  group_by(bmUnit) %>% arrange(halfHourEndTime) %>% 
  mutate(
    generation = quantity + lag(quantity)
  )

bmu_t0 %>% 
  summarise(quantity = sum(quantity))


hawkes_df <- readxl::read_xlsx(
  path = file.path(path,"gb_renewables_hawkes.xlsx"))


hawkes_df %>% 
  summarise(generation = sum(`Gen. (MW)`, na.rm = TRUE))
```

```{r}
bmu_t0 %>% 
  group_by(bmUnit) %>% arrange(halfHourEndTime) %>% 
  mutate(
    generation = quantity + lag(quantity)
  ) %>% ungroup() %>% 
  # filter(grepl("WLN",bmUnit) ) %>% 
  group_by(hour(halfHourEndTime)) %>% 
  summarise(quantity = sum(quantity), .groups = "drop")
```

```{r}
agpt25 %>% pull(startTime) %>% range()
agpt25 %>% 
  filter(
    startTime > as.POSIXct("2025-05-29 05:00:00", tz="UTC"),
    startTime <= as.POSIXct("2025-05-29 10:00:00", tz="UTC")) %>% 
  select(-businessType) %>% 
  pivot_wider(names_from = psrType, values_from = quantity) %>% 
  select(startTime:Solar) %>% 
  arrange(startTime) %>% view()
```

###check against energy trends report

#### 2024

```{r}
agpt24 <- fread(
  file.path(
    elexon_path,
    "data_by_year",
    "aggregated_gen_type_2024.csv.gz"))

agpt_tbl <- agpt24 %>% 
  # filter(
  #   grepl("Wind ",psrType)) %>% 
  group_by(startTime) %>% filter(publishTime == max(publishTime)) %>% ungroup() %>% 
  mutate(psrType = case_when(
    grepl("Wind", psrType) ~ "Wind", 
    grepl("Hydro", psrType) ~ "Hydro", 
    TRUE ~ psrType)) %>% 
  group_by(psrType) %>%
  arrange(psrType,startTime) %>% 
  mutate(generation = quantity+lag(quantity)) %>%
  filter(minute(startTime)==0) %>%
  group_by(psrType) %>% summarise(quantity = sum(quantity)/1e3) %>%
  filter(quantity>0) %>% 
  janitor::adorn_totals() 

agpt_tbl %>% 
  kable(digits = 0, col.names = c("Fuel type","Generation GWh"),
        format.args = list(big.mark = ",", scientific = FALSE)) %>% 
  kable_styling(full_width = F)



# agpt24 %>% 
#   filter(
#     grepl("Wind",psrType),
#     settlementDate %in% c("2024-12-18")
#   ) %>% 
#   group_by(startTime) %>% filter(publishTime == max(publishTime)) %>% 
#   select(psrType:settlementPeriod,matches("revision")) %>%
#   arrange(startTime) %>% #view()
#   pivot_wider(names_from = psrType, values_from = quantity) %>% 
#   # group_by(startTime) %>% 
#   # mutate(onshore= `Wind Onshore`+lag(`Wind Onshore`) ) %>%
#   mutate(total = `Wind Onshore` + `Wind Offshore`) %>% 
#   # ungroup() %>% 
#   filter(minute(startTime)==0)
```
```{r}
library(officer)
library(flextable)

ft <- agpt_tbl %>%
  mutate(quantity = scales::comma(quantity, accuracy = 1)) %>% 
  rename(`Fuel type` = psrType, `Generation GWh` = quantity) %>% 
  flextable() %>% 
  autofit() %>%
  align(align = "center", part = "all") %>%
  align(j = "Generation GWh", align = "right", part = "all") # right-align numbers

# Export to Word
read_docx() %>%
  body_add_flextable(ft) %>%
  print(target = "generation24_summary_fuel.docx")

```

```{r}
bmu_24 <- fread(
  file.path(path,"data_by_year","wind_gen_bmu_2024.csv.gz")
) %>% 
  filter(settlementDate %in% c("2024-12-18")) %>% 
  group_by(bmUnit) %>% arrange(halfHourEndTime) %>% 
  mutate(
    generation = quantity + lag(quantity)
  ) %>% ungroup() %>% 
  filter(
    halfHourEndTime > as.POSIXct("2024-12-18 03:00:00", tz="UTC"),
    halfHourEndTime <= as.POSIXct("2024-12-18 03:30:00", tz="UTC"),
    ) 

bmu_24 %>% 
  summarise(quantity = sum(generation))
```

#### 2023

```{r}
agpt23 <- fread(
  file.path(
    elexon_path,
    "data_by_year",
    "aggregated_gen_type_2022.csv.gz"))


agpt23 %>% 
  # filter(
  #   grepl("Wind ",psrType)) %>% 
  group_by(startTime) %>% filter(publishTime == max(publishTime)) %>% ungroup() %>% 
  group_by(psrType) %>%
  arrange(psrType,startTime) %>% 
  mutate(generation = quantity+lag(quantity)) %>%
  filter(minute(startTime)==0) %>%
  group_by(psrType) %>% summarise(
    quantity = sum(quantity)/1e3, generation = sum(generation, na.rm = T)/1e3) %>% 
  janitor::adorn_totals()
```

### number of units in Hawkes website

```{r}
sum(hawkes_df$`Gen. (MW)`!=0, na.rm = T)
```

### all units

```{r}
url0 <- "https://data.elexon.co.uk/bmrs/api/v1/datasets/B1610/stream?from=2024-12-18T00%3A00Z&to=2024-12-19T00%3A00Z"

bmu.string <- capacity %>% 
  filter(grepl("Wind|Solar",psrType, ignore.case = TRUE)  ) %>% 
  filter(!is.na(bmUnit)) %>% 
  pull(bmUnit) %>% 
  paste(.,collapse = "&bmUnit=")


# bmu.string <- "T_HOWBO-1&bmUnit=T_HOWBO-3&bmUnit=T_HOWBO-2&bmUnit=T_HOWAO-3&bmUnit=T_HOWAO-2&bmUnit=T_HOWAO-1"


url <- paste0(url0)

response <- GET(url, accept("text/plain"))
json_data <- content(response, "text", encoding = "UTF-8")
generation.bmu <- fromJSON(json_data) %>% 
  mutate(halfHourEndTime = as.POSIXct(halfHourEndTime, format = "%Y-%m-%dT%H:%M:%OS", tz = "UTC")) 
```

```{r}
generation.bmu %>% 
  group_by(halfHourEndTime) %>% 
  summarise(quantity = sum(quantity))

generation.bmu %>% pull(halfHourEndTime) %>% range()


generation.bmu$halfHourEndTime %>% tz()
generation.bmu %>% 
  filter(
    halfHourEndTime > as.POSIXct("2024-12-18 03:00", tz = "UTC"),
    halfHourEndTime <= as.POSIXct("2024-12-18 04:30", tz = "UTC")
    ) %>% 
  mutate(sign = sign(quantity)) %>% 
  group_by(halfHourEndTime, sign) %>% 
  summarise(quantity = sum(quantity))
```

### mapping catalogs

```{r}
wind.bmus
```

```{r}
repd_wind
```

```{r}
id_catalog %>% 
  filter(rep_id != "-")

bmu_to_dict <- id_catalog %>% 
  separate_rows(bmu_id, sep = ",\\s*") %>%
  separate_rows(rep_id, sep = ",\\s*") %>%
  select(dict_id, name, bmu_id, rep_id) %>% 
  unique() %>% 
  left_join(
    repd_subset %>% 
      select(ref_id:y_coord) %>% 
      mutate(
        ref_id = as.character(ref_id), 
        operator = `Operator (or Applicant)`,
        site_name = `Site Name`,
        tech_typ,
        capacity_repd = `Installed Capacity (MWelec)`),
    by = c("rep_id"="ref_id")
  )
```

```{r}
temp <- bmus %>% 
  filter(
    grepl("WIND",fuelType) | 
      grepl("wind|offshore", leadPartyName, ignore.case = TRUE )|
      grepl("wind farm|windfarm", bmUnitName, ignore.case = TRUE )) #%>%
  # filter(productionOrConsumptionFlag %in% c("P"))

temp %>% group_by(leadPartyName) %>% 
  summarise(capacity = sum(as.numeric(generationCapacity))) %>% 
  arrange(desc(capacity))
```

## generation outurn summary

```{r}
get_outturn <- function(
    year = 2024,
    path,
    file_name = sprintf("outturn_gen_type_%d.csv.gz",year),
    end_time = NULL
){
  
  if (is.null(end_time)){
    time_seq <- dates <- seq.Date(
    from = as.Date(sprintf("%d-01-01",year)),
    to = as.Date(sprintf("%d-12-31",year)),
    by = "4 days"
    )
    time_seq <- c(time_seq, as.Date(sprintf("%d-01-01",year+1)))
  } else{
    time_seq <- dates <- seq.Date(
    from = as.Date(sprintf("%d-01-01",year)),
    to = as.Date(end_time),
    by = "4 days"
    )
    if(time_seq[length(time_seq)]<end_time) time_seq <- c(time_seq,end_time)
  }
  
  
  # browser()
  extraction <- lapply(
    seq_along(time_seq[-1]),
    
    
    \(i){
      
    t0 <- as.character(time_seq[i])
    t1 <- as.character(time_seq[i+1])
    dt0 <- paste(t0, "01:00")
    dt1 <- paste(t1, "01:00")
    dt0_enc <- URLencode(dt0, reserved = TRUE)
    dt1_enc <- URLencode(dt1, reserved = TRUE)
    
    base_url <- "https://data.elexon.co.uk/bmrs/api/v1/generation/outturn/summary?"
    
    url <- sprintf(
      "%sstartTime=%s&endTimeTo=%s&includeNegativeGeneration=true&format=json",
      base_url,dt0_enc, dt1_enc)
    
    response <- GET(url, accept("text/plain"))
    json_data <- content(response, "text", encoding = "UTF-8")
    browser()
    blockdf <- json_data %>% 
      fromJSON(flatten = T) %>% 
      unnest(data) %>% 
      mutate(across(matches("Time"), ~ as.POSIXct(., format = "%Y-%m-%dT%H:%M:%OS")))
    blockdf}
  )
  
  df <- extraction %>% bind_rows()

  df %>% 
    write.csv(
      gzfile(file.path(path,file_name))
    )
  
  invisible(df)
}






outturn24 <- get_outturn(
  year = 2025,
  path = file.path(path,"data_by_year"),
  end_time = "2025-01-30")
# outturn24 <- outturn25
outturn24 <- get_outturn(
  year = 2025,
  path = file.path(path,"data_by_year"),
  end_time = "2025-05-30")
```

```{r}
 t0 <- "2024-01-01"
    # t1 <- as.character(time_seq[i+1])
    dt0 <- paste(t0, "01:00")
    # dt1 <- paste(t1, "01:00")
    dt0_enc <- URLencode(dt0, reserved = TRUE)
    # dt1_enc <- URLencode(dt1, reserved = TRUE)
    
    base_url <- "https://data.elexon.co.uk/bmrs/api/v1/generation/outturn/summary?"
    
    url <- sprintf(
      "%sstartTime=%s&includeNegativeGeneration=true&format=json",
      base_url,dt0_enc)
    
    response <- GET(url, accept("text/plain"))
    json_data <- content(response, "text", encoding = "UTF-8")
    # browser()
    blockdf <- json_data %>% 
      fromJSON(flatten = T) %>% 
      unnest(data) %>% 
      mutate(across(matches("Time"), ~ as.POSIXct(., format = "%Y-%m-%dT%H:%M:%OS")))
    blockdf
```

```{r}

```

```{r}
# blockdf %>% 
#   filter(startTime == c("2024-12-18"))
blockdf %>%
  mutate(year = year(startTime)) %>% 
  mutate(fuelType = ifelse(grepl("INT", fuelType),"INT",fuelType)) %>% 
  group_by(year, fuelType) %>% 
  summarise(
    n = n(),
    generation = sum(generation, na.rm = T)*24/1e6/n*365,
    .groups = "drop") %>% 
  select(-n) %>% 
  pivot_wider(names_from = fuelType, values_from = generation) %>% 
  view()
```

```{r}
agpt$psrType %>% unique() %>% paste(., collapse = ", ")
```

